# Analysis of TP53

## Data

### Discovery Sites

Three independent discovery studies focused on TP53 polymorphisms and risk of ovarian cancer: the North Carolina Ovarian Cancer Study (NCOCS), the Mayo Clinic Case-Control Study (MAYO), and the Polish Ovarian Cancer Study (POCS). These were restricted to non-Hispanic white women with newly diagnosed, histologically confirmed, primary invasive epithelial ovarian cancer and to non-Hispanic white controls. 23 SNPs were genotyped in total, with some overlap between sites.


### Replication Sites

Ten other sites contributed data: the Australian Ovarian Cancer Study (AOCS) and the Australian Cancer Study (ACS) presented together as AUS, the Family Registry for Ovarian Cancer (FROC, presented as STA), the Hawaiian Ovarian Cancer Study (HAW), the Malignant Ovarian Cancer Study Denmark (MALOVA), the New England Case-Control Study (NEC), the Nurses' Health Study (NHS), SEARCH Cambridge (SEA), the Los Angeles County Case-Control Study of Ovarian Cancer (LAC-CCOC, presented here as USC), the University of California at Irvine study (UCI), and the United Kingdom Ovarian Cancer Population Study (UKOPS, presented here as UKO).

The combined data set (discovery and replication) comprised 5,206 white, non-Hispanic invasive epithelial ovarian cancer cases, of which 2,829 were classified as serous invasive ovarian cancer, and 8,790 white non-Hispanic controls. Analysis was restricted to white, non-Hispanic invasive serous ovarian cancer cases and white, non-Hispanic controls.

## Models

Each model adjusts for study site, reference age, and personal history of breast cancer. History of breast cancer is treated as a fixed effect, and the rest of the covariates's coefficients are treated as normally distributed random effects. 

$$P(Y_{ij}=1) = \textsf{logit}^{-1}(\beta^{site}_{ j}+ \beta^{p53}_{j}*p53_{ij}+ \beta^{age}_{j}*age_{ij}+\beta^{BC}* BC_{ij})$$

$$\beta^{site}_{ j}\sim \textsf{N}(\mu^{site}, \sigma^{2site})$$
$$\beta^{age}_{j}\sim \textsf{N}(\mu^{age}, \sigma^{2age})$$
$$\beta^{BC}, \mu^{age},\mu^{site}\sim \textsf{N}(0,0.1)$$
$$\sigma^{age},\sigma^{site} \sim \textsf{invGamma}(1, 0.05)$$

The p53 variable's site-specific log OR priors were defined using the models described previously.

The fully Bayesian model was fit jointly as well as marginally. Since the results were very similar, the marginal models were used for computational efficiency and clarity of interpretation. Results from the joint analysis can be found in the supplement.


## Results

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(cache = TRUE)

library(rmeta)
library(lme4)
library(R2jags)
library(xtable)
library(knitr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(ggthemes)

source("HPD.R")
source("p53finalJAGScode.R")
load("p53OR.RData")
load("tp53.RData")
```


```{r load data}
getdata<- function(snp.name, iter=5000, drop.sites= NULL){
  load("tp53.Rdata")
  if(!is.null(drop.sites)) {
    
    use = !(tp53epi.wsi$site %in% drop.sites)
  } else {
    use = rep(TRUE, nrow(tp53epi.wsi))
  }
  #this is a new indicator
  site.names = levels(factor(tp53epi.wsi[which(!is.na(tp53geno.wsi[,snp.name])),"site"]))
  discovery.sites <- which(site.names%in% drop.sites)
  
  p53.snp = tp53geno.wsi[use, snp.name]
  tp53epi.wsi = tp53epi.wsi[use,]
  
  missing.geno = is.na(p53.snp)
  site.names = levels(factor(tp53epi.wsi[!missing.geno,"site"]))
  
  p53.data = list(CaseCon=tp53epi.wsi$casecon[!missing.geno], site=as.numeric(factor(tp53epi.wsi$site[!missing.geno])), BC = as.numeric(tp53epi.wsi$prev.BC[!missing.geno]), Age = tp53epi.wsi$refage[!missing.geno], p53 = p53.snp[!missing.geno])
  
  p53.df = data.frame(p53.data)
  
  J = length(p53.data$CaseCon)
  n.sites = length(unique(p53.data$site))
  p53.data$J = J
  p53.data$n.sites = n.sites
  
  p53.data$discovery.sites<-discovery.sites
  p53.data$missing.geno<-missing.geno
  p53.data$site.names<-site.names
  
  return(p53.data)
  
}

```

```{r frequentist analysis function}
OR.freq = function(snp.name, psdir="ps", drop.sites=NULL,iter=5000, debug=F) {
  load("tp53.Rdata")
  if (!is.null(drop.sites)) {
    use = !(tp53epi.wsi$site %in% drop.sites)}
  else {
    use = rep(TRUE, nrow(tp53epi.wsi))
  }
  
  p53.snp = tp53geno.wsi[use, snp.name]
  tp53epi.wsi = tp53epi.wsi[use,]
  
  missing.geno = is.na(p53.snp)
  site.names = levels(factor(tp53epi.wsi[!missing.geno,"site"]))
  
  p53.data = list(CaseCon=tp53epi.wsi$casecon[!missing.geno], site=as.numeric(factor(tp53epi.wsi$site[!missing.geno])), BC = as.numeric(tp53epi.wsi$prev.BC[!missing.geno]), Age = tp53epi.wsi$refage[!missing.geno], p53 = p53.snp[!missing.geno])
  
  p53.df = data.frame(p53.data)
  write.csv(p53.df, file=paste(snp.name, ".csv", sep=""))
  if(length(site.names)>1){
    p53.full = glm(CaseCon ~ factor(site) + factor(site)*p53 + factor(site)*Age + BC, data=p53.df, family=binomial, x=T)
    p53.pooled = glm(CaseCon ~ factor(site) + p53 + factor(site)*Age + BC, data=p53.df, family=binomial)
    p53.null = glm(CaseCon ~ factor(site) +  factor(site)*Age + BC, data=p53.df, family=binomial)
    
    p53.df$Site = factor(p53.df$site)
    
    p53.me = glmer(CaseCon ~ BC + p53 + Age + (1|site)  + (0 + p53 | site) + (0 + Age | site), start=c(site=1.0, p53=.1, Age=.01),nAGQ=1 , data=p53.df, family=binomial)
    
    test=anova(p53.full, p53.pooled, p53.null, test="Chi")
    coef = summary(p53.full)$coef
    ns = length(site.names)
    OR = coef[c(ns+1, (ns+4):(ns+ns+2)),1]
    x = p53.full$x
    p = predict(p53.full, type="response")
    var = solve(t(x)%*% diag(p*(1-p)) %*%x)[c(ns+1, (ns+4):(ns+ns+2)),c(ns+1, (ns+4):(ns+ns+2))]
    
    sqrt(diag(var))
    
    eff = matrix(0, ns,ns)
    eff[,1] = 1
    for (i in 2:ns) eff[i,i] = 1
    
    OR = eff %*% OR
    OR.SE = sqrt(diag(eff %*% var %*% t(eff)))
    
    DS = meta.summaries(OR, OR.SE, method="random", names=site.names, logscale=F)
    p.value = pnorm(-(abs(DS$summary/DS$se.summary)))*2
    BF0 = -exp(1)*p.value*log(p.value)
    return(list(snp=snp.name, DS=DS, OR=OR, SE=OR.SE, p.value=p.value, BF.Ha = 1/BF0, test=test, p53.me=p53.me))}
  else{
    p53.full = glm(CaseCon ~ p53 + Age + BC, data=p53.df, family=binomial, x=T)
    p53.null = glm(CaseCon ~ Age + BC, data=p53.df, family=binomial)
    test=anova(p53.full, p53.null, test="Chi")
    coef = summary(p53.full)$coef
    ns = length(site.names)
    OR = as.matrix(coef[2,1])
    OR.SE = (coef[2,2])
    p.value = coef[2,4]
    BF0 = -exp(1)*p.value*log(p.value)
    return(list(snp=snp.name, DS=DS, OR=OR, SE=OR.SE, p.value=p.value, BF.Ha = 1/BF0))
  }
}

```


```{r comparison}
getOR<-function(sim, totalsites,discoverysites){
  
  OR = as.data.frame(sim$BUGSoutput$sims.matrix)%>%select( starts_with("beta.p53"), "mu.p53")
  exp(OR)
  #colnames(OR) = c(site.names, "Overall")
  
  sum.OR = t(apply(exp(OR), 2, function(x) {PI = HPDM(as.matrix(x),pointmass=1)
  return(c(median(x), PI[1], PI[2], PI[3],PI[4]))} #how to deal w multi, change HPD to have point mass at 1
  ))
  if (dim(sum.OR)[1]!= totalsites+1){
    sum.OR1<- matrix(NA, totalsites+1, dim(sum.OR)[2])
    sum.OR1[-discoverysites,]<- sum.OR
    sum.OR<- sum.OR1
  } ###
  #colnames(sum.OR)<-c("Median","2.5%","97.5%")
  return(sum.OR)
}

```



```{r run all p53}
OR.runall<-function(snp.name, discovery.sitenames, iter=5000,p = 0.00325, psdir="ps"){
  #data
  
  p53.data<- getdata(snp.name, drop.sites=NULL)
  p53.dataval<- getdata(snp.name, drop.sites=discovery.sitenames)
  validation.sitenames = setdiff(p53.data$site.names, discovery.sitenames)
  freq<-OR.freq(snp.name, psdir="ps",
                drop=validation.sitenames,iter=5000, debug=F) 
  p53.data.normal<-p53.dataval
  p53.data.normal$n.discovery<- length(p53.data.normal$discovery.sites)
  p53.data.normal$zeroes<- rep(0,p53.data.normal$n.discovery)
  p53.data.normal$MLE<- freq$OR[,1] 
  p53.data.normal$SE<- freq$SE 
  p53.data.normal$q<-qnorm(1-p/2)
  bfdata <- p53.dataval
  bfdata$p <- freq$p.value
  
  #run all 
  parameters = c("beta.BC", "beta.Age", "mu.site","mu.p53", "mu.Age", "sigma.site", "sigma.p53",  "sigma.Age" ,"assoc", "pind","beta.site", "beta.p53","phi.site","phi.p53")
  
  p53.sim = jags(data=p53.data, inits=NULL, parameters.to.save =parameters, model = p53.model)
  p53.simnew = jags(data=p53.data , inits=NULL, parameters.to.save =parameters, model = p53.newmodel) 
  p53.simnormal = jags(data=p53.data.normal, inits=NULL, parameters.to.save =parameters.normal, model = p53.normal)
  p53.bfsim = jags(data=bfdata, inits=function(){list("pind"=.9)}, parameters.to.save =parameters, model = p53.bf.approx)
  
  #OR
  return(list(original= (p53.sim),
              bayes= (p53.simnew),
              cond= (p53.simnormal),
              bf= (p53.bfsim),
              n.sites = p53.data$n.sites,
              n.discovery = p53.data.normal$discovery.sites))
}


```




```{r, eval= FALSE}
discovery.siteslist<-list(c("POL"),
                          c("POL", "MAY", "NCO"),
                          c("POL"),
                          c("POL"), 
                          c("NCO"),
                          c("POL", "MAY", "NCO"),
                          c("NCO")
                          
) 

snps<- setdiff(unique(colnames(tp53geno.wsi)), c( "rs2909430n", "rs2287499n", "rs2078486n"))

allSNP<- vector("list", length = length(snps))
for(i in 1:length(snps)){
  allSNP[[i]]<- OR.runall(snps[i],discovery.sitenames= discovery.siteslist[[i]])
  
}
save.image()

ORbySNP2<- lapply(allSNP, function(snplist)
  lapply(snplist[1:(length(snplist)-2)],function(x)
    getOR(x,snplist[[(length(snplist)-1)]],snplist[[(length(snplist))]])))
```


```{r}
snps<- setdiff(unique(colnames(tp53geno.wsi)), c( "rs2909430n", "rs2287499n", "rs2078486n"))[c(1,2,3,4,6)] #only the ones needed
ORbySNP<-ORbySNP[c(1,2,3,4,6)]

overall = do.call("rbind", lapply(ORbySNP, function(y) data.frame(t(sapply(y, function(x) (x[dim(x)[1],]))))))

#overall=data.frame(t(sapply(ORbySNP[[2]], function(x) x[dim(x)[1],])))
colnames(overall) = c("X","L","U", "L1","U1")
overall$model = rep(c("original","bayesian","conditional","bf approx"),length(snps))
overall$model<- factor(overall$model, levels= c("bf approx", "conditional","bayesian","original"))
overall$snp = rep(snps, each = 4)
ggplot(data = overall, aes(group=model,color=model,x = snp,xend = snp))+ 
  geom_hline(yintercept=1, alpha=.3)+
  geom_linerange(aes(ymin=L, ymax=U), position = position_dodge(width = .7))+
  geom_linerange(aes(ymin=L1, ymax=U1), position = position_dodge(width = .7))+ 
  geom_point(aes(x=snp, y=X),size=3, shape=4,position = position_dodge(width = .7)) +
  coord_flip()+theme_minimal() + scale_colour_few()+labs(y="Odds Ratio", title= "OR and 95% CI for p53 SNPs")
  

```



The new models shrunk estimates towards 0 for all SNPs. The BF approximation model set all point estimates to zero, as did the fully Bayesian model with the exception of one. This is the same SNP (`r overall$snp[which(overall$X!=1 &overall$model=="bayesian")]`) that was detected as significant in the MISA analysis [@schildkraut2010association]. However, the credible interval contains 0, so there is not enough evidence for association. Similarly, although the conditional likelihood model does not set any estimates to zero, all its credible intervals contain zero.


The original model has the longest credible intervals for all SNPs except `r overall$snp[which(overall$X!=1 &overall$model=="bayesian")]`, for which both the Bayesian and Bayes Factor models have large credible intervals, indicating more uncertainty than the original model.

Below are the full tables of OR by site and SNP.

> Should I add some plots of posteriors of mu or the beta's?
> Should I have some summary of the other covariates (age, site, etc)?
> Would it be better to put these tables elsewhere?


```{r, results="asis"}
all.sitenames<-lapply(snps, function(x) getdata(x)$site.names)

formatCI<- function(ORvec){
  if(is.na(ORvec[2])) ret<-"-"
  else ret<-paste(ORvec[2],"-",ORvec[3])
  if (!is.na(ORvec[4])){
    ret<-paste(ret, ", ", ORvec[4],"-",ORvec[5])
  }
  return(ret)
}
makeORtable<-function(OR, sitenames){
  est<-sapply(OR, function(x) paste(round(x[,1], 3)))
  CI<-sapply(OR, function(x) apply(round(x, 3),1, formatCI))
  ret<-matrix(NA, nrow = dim(est)[1],ncol=8)
  ret[,-c(2,4,6,8)]<-est
  ret[,c(2,4,6,8)]<-CI
  colnames(ret)<- c("original estimate","original CI",
                    "bayesian estimate","bayesian CI",
                    "CL estimate","CL CI",
                    "BF estimate","BF CI")
  rownames(ret)<- c(sitenames, "Overall")
  return(ret)
}

for(i in 1:length(snps)){
  print(kable(makeORtable(ORbySNP[[i]], all.sitenames[[i]]), caption=snps[i]))
}
```
